Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2021-09-15 paper/paper.tex
update benchmark script

\begin{table}[t!]\centering
\begin{minipage}{\columnwidth}
\ra{1.3}
        \begin{tabularx}{\textwidth}{sb}\toprule
            \hline
            \textbf{element type}     & \textbf{purpose} \\
            {$\mathbbm{R}$}     & {real number for counting all independent sets} \\
            {$\mathbbm{C}$}     & {complex number for fitting the independence polynomial with fast Fourier transform} \\
            {$\text{GF}(p)$} (\Eq{eq:finitefield}) & {number on finite field algebra for fitting the independence polynomial exactly using number theory} \\
            {T} (\Eq{eq:tropical})    & {tropical number for finding the independence number} \\
            {P1} (\Eq{eq:countingtropical})     & {polynomial truncated to largest order, i.e. the tropical number with counting field for finding the independence number and the number of MISs} \\
            {P2} (\Eq{eq:max2poly})     & {counting MISs and independent sets of size $\alpha(G)-1$} \\
            {PN} (\Eq{eq:polynomial})     & {computing the independence polynomial exactly} \\
            {P1+S1}     & {replacing the counting field in P1 to configuration sampler defined in $\Eq{eq:singleconfig}$ for finding the independence number and one MIS} \\
            {P1+SN} (\Eq{eq:countingtropicalset})    & {replacing the counting field in P1 to configuration sampler defined in $\Eq{eq:set}$ for finding the independence number and enumeration of all MISs} \\
            \bottomrule
        \end{tabularx}
    \caption{Tensor element types and their purposes in calculating various independent set properties.}\label{tbl:generictypes}
\end{minipage}
\red{we should include maximal independence polynomial and counting, enumeration in this table as well.}
\end{table}

@@

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets on a random three regular graph with different tensor element types. \red{for each graph size, is there only one graph?}\blue{yes} (a) treewidth versus the number of vertices for the benchmarked graphs. 
    (b) The computing time for calculating the independence number and for counting of the number of independent sets, the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$.
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for enumeration problems, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    \red{suggestions for the figure: 1. the legend seems to overlap the lines, especially the large subfigure. Not sure what's a better way to include the legend there. 2. I see you guys don't typically use markers. We typically use markers to distinguish the different lines, as well as using the color. Especially, for some journals, there is a physical non-color copy, so some requires the figures to be understandable without colors.}
    }
    \label{fig:benchmark}
\end{figure}

The computations shown in \Fig{fig:benchmark} include \texttt{max size} (independence number, $\alpha(G)$), \red{we can consider changing `max size' to `ind number', but I am okay with both}
 \texttt{total counting} (total number of independent sets) \red{should we change total size to total counting here, since max size meant max size of the set}, 
 \texttt{max counting} (counting of MISs), 
 \texttt{max2 counting} (counting of independent sets having a size $\alpha(G)$ or $\alpha(G)-1$), \red{we can consider changing `max counting' and `max2 counting' to `MIS counting' and `MIS and MIS-1 counting', but I am okay with both notations}
 \texttt{IDP} (independence polynomial), 
 \texttt{single MIS} (find one MIS configuration), 
 \texttt{single MIS (bounding)} (find one MIS configuration including the bounding of enumeration space), \texttt{all MIS} (enumerate all MIS configurations), 
\texttt{all MIS (bounding)} (enumerate all MIS configurations including the bounding of enumeration space),
\texttt{all MIS and MIS-1} (enumeration of all independent sets having a size $\alpha(G)$ or $\alpha(G)-1$),
\texttt{all IS} (enumeration of all independent sets).
\red{Jinguo, should we include maximal independence polynomial and enumeration of all maximal independent sets as well? I think that will make the benchmark more complete.}

Among the benchmarks, computing \texttt{max size (CPU)},
\texttt{total size (CPU)}, \texttt{total size (GPU)}, \texttt{IDP (FFT) (CPU)}, \texttt{IDP (FFT) (GPU)} and \texttt{single configuration (bounding) (CPU)} can utilize fast BLAS functions, and hence 
they are faster comparing to non-BLAS methods in the same category.
Only algebras requiring dynamic sized structures cannot be computed on GPU, which includes the polynomial algebra for computing the independence polynomial and the set algebra for enumerating configurations.
When computing the independence polynomial, the finite field approach is the only method that does not incur round-off errors and that can be run on a GPU.
To enumerate the MISs, the technique to bound the enumeration space is indispensible,
without which the largest enumerable size is about $150$ vertices with 32G memory while also being one order-of-magnitude slower.
All these computing times have a strong correlation with the treewidth, 
and the computing times of configuration enumeration also depend on the size of the target configuration space.

@@

We also benchmarked our method with the Bron-Kerbosch method to enumerate and count maximal independent sets in \Fig{fig:benchmark-maximal}.
We can see the tree width of this tensor network is nearly doubled comparing with the one for computing independent sets, hence only small graphs can be benchmarked.
The time for the tensor netwrok approach and Bron-Kerbosch approach to enumerate all maximal independent sets are comparable,
while the counting can be much faster.

>>>>

\begin{table}[t!]\centering
\begin{minipage}{\columnwidth}
\ra{1.3}
        \begin{tabularx}{\textwidth}{sb}\toprule
            \hline
            \textbf{element type}     & \textbf{purpose} \\
            {$\mathbbm{R}$}     & {real number for counting all independent sets} \\
            {$\mathbbm{C}$}     & {complex number for fitting the independence polynomial with fast Fourier transform} \\
            {$\text{GF}(p)$} (\Eq{eq:finitefield}) & {number on finite field algebra for fitting the independence polynomial exactly using number theory} \\
            {T} (\Eq{eq:tropical})    & {tropical number for finding the independence number} \\
            {P1} (\Eq{eq:countingtropical})     & {polynomial truncated to largest order, i.e. the tropical number with counting field for finding the independence number and the number of MISs} \\
            {P2} (\Eq{eq:max2poly})     & {counting MISs and independent sets of size $\alpha(G)-1$} \\
            {PN} (\Eq{eq:polynomial})     & {computing the independence polynomial exactly} \\
            {P1+S1}     & {replacing the counting field in P1 to configuration sampler defined in $\Eq{eq:singleconfig}$ for finding the independence number and one MIS} \\
            {P1+SN} (\Eq{eq:countingtropicalset})    & {replacing the counting field in P1 to configuration sampler defined in $\Eq{eq:set}$ for finding the independence number and enumeration of all MISs} \\
            \bottomrule
        \end{tabularx}
    \caption{Tensor element types and their purposes in calculating various independent set properties.}\label{tbl:generictypes}
\end{minipage}
\red{we should include maximal independence polynomial and counting, enumeration in this table as well.}
\blue{I do not think so. Network structure and element types are two degree of freedoms,
maximal IS is different network structure. We can use the above algebra in any network structure.}
\end{table}

@@

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets on a random three regular graph with different tensor element types. \red{for each graph size, is there only one graph?}\blue{yes} (a) treewidth versus the number of vertices for the benchmarked graphs. 
    Legends are the data types that we used in the computation, one can find the corresponding computational task in \Tbl{tbl:generictypes}.
    (b) The computing time for calculating the independence number and for counting of the number of independent sets, the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$.
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for enumeration problems, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    \red{suggestions for the figure: 1. the legend seems to overlap the lines, especially the large subfigure. Not sure what's a better way to include the legend there. 2. I see you guys don't typically use markers. We typically use markers to distinguish the different lines, as well as using the color. Especially, for some journals, there is a physical non-color copy, so some requires the figures to be understandable without colors.}
    \blue{If we use markers, the data poitns are too dense. I prefer not adding markers, not much people read papers with kindle or on paper.}
    \blue{I used a smaller fontsize, there are still some overlaps. Does it look better (or worse) now?}
    }
    \label{fig:benchmark}
\end{figure}

The computations shown in \Fig{fig:benchmark} include independent set counting, finding maximum independent set size (panel b),
computing independence polynomial (panel c) and enumerating and sampling independent sets (panel d).
\blue{I cleaned up this paragraph, because one can check the table \Tbl{tbl:generictypes} now.}
\red{Jinguo, should we include maximal independence polynomial and enumeration of all maximal independent sets as well? I think that will make the benchmark more complete.}
\blue{added a new paragraph.}
Among the benchmarks, computing \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions, and hence 
they are faster comparing to non-BLAS methods in the same category.
Only algebras requiring dynamic sized structures cannot be computed on GPU, which includes \texttt{PN} and \texttt{SN}.
When computing the independence polynomial, the finite field (GF$(p)$) approach is the only method that does not incur round-off errors and that can be run on a GPU.
To enumerate the MISs, the technique to bound the enumeration space is indispensible,
without which the largest enumerable size is about $150$ vertices with 32G memory while also being one order-of-magnitude slower.
All these computing times have a strong correlation with the treewidth, 
and the computing times of configuration enumeration also depend on the size of the target configuration space.

@@

We also benchmarked our method with the Bron-Kerbosch method to enumerate and count maximal independent sets in \Fig{fig:benchmark-maximal}.
Since the network for computing maximal independent set properties is different from the one for computing the independence polynomial,
the tree width of this tensor network is significantly larger, hence only small graphs can be benchmarked.
The time for the tensor netwrok approach and the Bron-Kerbosch approach to enumerate all maximal independent sets are comparable,
while the tensor network does counting much more efficiently.
The implementation of Bron-Kerbosch algorithm is from the package LightGraphs~\cite{LightGraphs},
due to the memory limit, this algorithm stops working at size $70$ and above.
